---
title: "Correlation and regression fundamentals with tidy data principles"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# INTRO ----

While the tidymodels package broom is useful for summarizing the result of a single analysis in a consistent format,
it is really designed for high-throughput applications, where you must combine results from multiple analyses.
These could be subgroups of data, analyses using different models, bootstrap replicates, permutations, and so on.
In particular, it plays well with the nest()/unnest() functions from tidyr and the map() function in purrr.


# CORRELATION ANALYSIS ----

```{r message=FALSE, warning=FALSE}
library(tidymodels)
library(tidyverse)

orange <- as_tibble(Orange)

orange
```

This contains 35 observations of three variables: Tree, age, and circumference.
Tree is a factor with five levels describing five trees. As might be expected, age and circumference are correlated:

```{r}
cor(orange$age, orange$circumference)

orange %>% 
    ggplot(aes(age, circumference, color = Tree)) +
    geom_line(size = 2)
```

Suppose you want to test for correlations individually within each tree. You can do this with dplyr’s group_by:

```{r}
orange %>% 
    group_by(Tree) %>% 
    summarise(correlation = cor(age, circumference))
```


(Note that the correlations are much higher than the aggregated one, and also we can now see the correlation is similar across trees).

Suppose that instead of simply estimating a correlation, we want to perform a hypothesis test with cor.test():

```{r}
ct <- cor.test(orange$age, orange$circumference)

ct
```


This test output contains multiple values we may be interested in. Some are vectors of length 1, such as the p-value and the estimate,
and some are longer, such as the confidence interval. We can get this into a nicely organized tibble using the tidy() function:

```{r}
tidy(ct)
```


Often, we want to perform multiple tests or fit multiple models, each on a different part of the data.
In this case, we recommend a nest-map-unnest workflow. For example, suppose we want to perform correlation tests for each different tree.
We start by nesting our data based on the group of interest:

```{r}
nested <- orange %>% 
    nest(data = c(age, circumference))
```


Then we perform a correlation test for each nested tibble using purrr::map():

```{r}
nested %>% 
    mutate(test = map(data, ~ cor.test(.x$age, .x$circumference)))
```

This results in a list-column of S3 objects. We want to tidy each of the objects, which we can also do with map()

```{r}
nested %>% 
    mutate(test = map(data, ~ cor.test(.x$age, .x$circumference)), # S3 list-col
           tidied = map(test, tidy))
```

Finally, we want to unnest the tidied data frames so we can see the results in a flat tibble. All together, this looks like:

```{r}
orange %>% 
    nest(data = c(age, circumference)) %>% 
    mutate(
        test = map(data, ~ cor.test(.x$age, .x$circumference)), # S3 list-col
        tidied = map (test, tidy)
    ) %>% 
    unnest(cols = tidied) %>% 
    select(-c(data, test))
```


# REGRESSION MODELS ----

This type of workflow becomes even more useful when applied to regressions. Untidy output for a regression looks like:

```{r}
lm_fit <- lm(age ~ circumference, data = orange)

summary(lm_fit)
```


When we tidy these results, we get multiple rows of output for each model:

```{r}
tidy(lm_fit)
```

Now we can handle multiple regressions at once using exactly the same workflow as before:

```{r}
orange %>% 
    nest(data = c(-Tree)) %>% 
    mutate(
        fit = map(data, ~ lm(age ~ circumference, data = .x)),
        tidied = map(fit, tidy)
    ) %>% 
    unnest(tidied) %>% 
    select(-data, -fit)
```

You can just as easily use multiple predictors in the regressions, as shown here on the mtcars dataset.
We nest the data into automatic vs. manual cars (the am column), then perform the regression within each nested tibble.

```{r}
mtcars

mtcars <- as_tibble(mtcars)

mtcars

mtcars %>%
    nest(data = c(-am)) %>% 
    mutate(
        fit = map(data, ~ lm(wt ~ mpg + qsec + gear, data = .x)),  # S3 list-col
        tidied = map(fit, tidy)
    ) %>% 
    unnest(tidied) %>% 
    select(-data, -fit)
```


What if you want not just the tidy() output, but the augment() and glance() outputs as well,
while still performing each regression only once? Since we’re using list-columns,
we can just fit the model once and use multiple list-columns to store the tidied, glanced and augmented outputs.

```{r}
regressions <- mtcars %>% 
    nest(data = c(-am)) %>% 
    mutate(
        fit = map(data, ~ lm(wt ~ mpg + qsec + gear, data = .x)),
        tidied = map(fit, tidy),
        glanced = map(fit, glance),
        augmented = map(fit, augment)
    )
```

```{r}
regressions %>% 
    select(tidied) %>% 
    unnest(tidied)
```

```{r}
regressions %>% 
    select(glanced) %>% 
    unnest(glanced)
```

```{r}
regressions %>% 
    select(augmented) %>% 
    unnest(augmented)
```

By combining the estimates and p-values across all groups into the same tidy data frame (instead of a list of output model objects),
a new class of analyses and visualizations becomes straightforward. This includes:

- sorting by p-value or estimate to find the most significant terms across all tests,

- p-value histograms, and

- volcano plots comparing p-values to effect size estimates

In each of these cases, we can easily filter, facet, or distinguish based on the term column.
In short, this makes the tools of tidy data analysis available for the results of data analysis and models, not just the inputs.
