---
title: "Hypothesis testing using resampling and tidy and data"
author: "VC"
date: "4/2/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# INTRODUCTION

The tidymodels package infer implements an expressive grammar to perform statistical inference that coheres with the tidyverse design framework. Rather than providing methods for specific statistical tests, this package consolidates the principles that are shared among common hypothesis tests into a set of 4 main verbs (functions), supplemented with many utilities to visualize and extract information from their outputs.

Regardless of which hypothesis test we’re using, we’re still asking the same kind of question:

    Is the effect or difference in our observed data real, or due to chance?

To answer this question, we start by assuming that the observed data came from some world where “nothing is going on” (i.e. the observed effect was simply due to random chance), and call this assumption our **null hypothesis**. (In reality, we might not believe in the null hypothesis at all; the null hypothesis is in opposition to the **alternate hypothesis**, which supposes that the effect present in the observed data is actually due to the fact that “something is going on.") We then calculate a **test statistic** from our data that describes the observed effect. We can use this test statistic to calculate a **p-value**, giving the probability that our observed data could come about if the null hypothesis was true. If this probability is below some pre-defined **significance level α**, then we can reject our null hypothesis.

If you are new to hypothesis testing, take a look at

- Section 9.2 of Statistical Inference via Data Science (https://moderndive.com/9-hypothesis-testing.html#understanding-ht)
- The American Statistical Association’s recent statement on p-values (https://www.tandfonline.com/doi/full/10.1080/00031305.2016.1154108)

The workflow of this package is designed around these ideas. Starting from some data set,

- specify() allows you to specify the variable, or relationship between variables, that you’re interested in,

- hypothesize() allows you to declare the null hypothesis,

- generate() allows you to generate data reflecting the null hypothesis, and

- calculate() allows you to calculate a distribution of statistics from the generated data to form the null distribution.

Throughout this vignette, we make use of gss, a data set available in infer containing a sample of 500 observations of 11 variables from the General Social Survey.

```{r}
library(tidymodels) # includes the infer package

gss %>% glimpse()

?gss

```

Each row is an individual survey response, containing some basic demographic information on the respondent as well as some additional variables. See ?gss for more information on the variables included and their source. Note that this data (and our examples on it) are for demonstration purposes only, and will not necessarily provide accurate estimates unless weighted properly. For these examples, let’s suppose that this data set is a representative sample of a population we want to learn about: American adults.


# SPECIFY VARIABLES

The specify() function can be used to specify which of the variables in the data set you’re interested in. If you’re only interested in, say, the age of the respondents, you might write:

```{r}
gss %>% 
    specify(response = age)
```

On the front end, the output of specify() just looks like it selects off the columns in the dataframe that you’ve specified. What do we see if we check the class of this object, though?

```{r}
gss %>% 
    specify(response = age) %>% 
    class()
```

We can see that the infer class has been appended on top of the dataframe classes; this new class stores some extra metadata.

If you’re interested in two variables (age and partyid, for example) you can specify() their relationship in one of two (equivalent) ways:

```{r}
gss %>% 
    specify(age ~ partyid) # as formula
```

```{r}
gss %>% 
    specify(response = age, explanatory = partyid) # with the named arguments
```

If you’re doing inference on one proportion or a difference in proportions, you will need to use the success argument to specify which level of your response variable is a success. For instance, if you’re interested in the proportion of the population with a college degree, you might use the following code:

```{r}
gss %>% 
    specify(response = college, success = "degree") # specify for inference on proportions
```


# DECLARE THE HYPOTHESIS

The next step in the infer pipeline is often to declare a null hypothesis using hypothesize(). The first step is to supply one of “independence” or “point” to the null argument. If your null hypothesis assumes independence between two variables, then this is all you need to supply to hypothesize():

```{r}
gss %>% 
    specify(college ~ partyid, success = "degree") %>% 
    hypothesise(null = "independence")
```

If you’re doing inference on a point estimate, you will also need to provide one of p (the true proportion of successes, between 0 and 1), mu (the true mean), med (the true median), or sigma (the true standard deviation). For instance, if the null hypothesis is that the mean number of hours worked per week in our population is 40, we would write:

```{r}
gss %>% 
    specify(response = hours) %>% 
    hypothesise(null = "point", mu = 40)
```

Again, from the front-end, the dataframe outputted from hypothesize() looks almost exactly the same as it did when it came out of specify(), but infer now “knows” your null hypothesis.


# GENERATE THE DISTRIBUTION

Once we’ve asserted our null hypothesis using hypothesize(), we can construct a null distribution based on this hypothesis. We can do this using one of several methods, supplied in the type argument:

- bootstrap: A bootstrap sample will be drawn for each replicate, where a sample of size equal to the input sample size is drawn (with replacement) from the input sample data.

- permute: For each replicate, each input value will be randomly reassigned (without replacement) to a new output value in the sample.

- simulate: A value will be sampled from a theoretical distribution with parameters specified in hypothesize() for each replicate. (This option is currently only applicable for testing point estimates.)

Continuing on with our example above, about the average number of hours worked a week, we might write:

```{r}
gss %>% 
    specify(response = hours) %>% 
    hypothesise(null = "point", mu = 40) %>% 
    generate(reps = 5000, type = "bootstrap")
```

In the above example, we take 5000 bootstrap samples to form our null distribution.

To generate a null distribution for the independence of two variables, we could also randomly reshuffle the pairings of explanatory and response variables to break any existing association. For instance, to generate 5000 replicates that can be used to create a null distribution under the assumption that political party affiliation is not affected by age:

```{r}
gss %>% 
    specify(partyid ~ age) %>% 
    hypothesise(null = "independence") %>% 
    generate(reps = 5000, type = "permute")
```


# CALCULATE STATISTICS

Depending on whether you’re carrying out computation-based inference or theory-based inference, you will either supply calculate() with the output of generate() or hypothesize(), respectively. The function, for one, takes in a stat argument, which is currently one of "mean", "median", "sum", "sd", "prop", "count", "diff in means", "diff in medians", "diff in props", "Chisq", "F", "t", "z", "slope", or "correlation". For example, continuing our example above to calculate the null distribution of mean hours worked per week:

```{r}
gss %>% 
    specify(response = hours) %>% 
    hypothesise(null = "point", mu = 40) %>% 
    generate(reps = 5000, type = "bootstrap") %>% 
    calculate(stat = "mean")
```

The output of calculate() here shows us the sample statistic (in this case, the mean) for each of our 1000 replicates. If you’re carrying out inference on differences in means, medians, or proportions, or t and z statistics, you will need to supply an order argument, giving the order in which the explanatory variables should be subtracted. For instance, to find the difference in mean age of those that have a college degree and those that don’t, we might write:

```{r}
gss %>%
    specify(age ~ college) %>%
    hypothesize(null = "independence") %>%
    generate(reps = 5000, type = "permute") %>%
    calculate("diff in means", order = c("degree", "no degree"))

```


# OTHER UTILITIES

